# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
myData <- read.csv("activity.csv")
myProcData <- myData[is.na(myData$steps)=="FALSE",]
```

## What is mean total number of steps taken per day?
```{r}
ii <- 2:length(myProcData$steps)
tempDate <- myProcData$date[1]
tempSum <- myProcData$steps[1]
totalSum <- matrix()
j <- 1
for (i in ii)
{
  currentDate <- myProcData$date[i]
  if(currentDate == tempDate)
  {
    tempSum <- tempSum + myProcData$steps[i]
  }
  else
  {
      totalSum[j] <- tempSum
      j <- j + 1
      tempSum <- myProcData$steps[i]
      tempDate <- myProcData$date[i]
  }
}
totalSum[j] <- tempSum ## total sum of the last analyzed data
```

```{r fig.width=7, fig.height=6}
hist(totalSum) ## Histogram of the total number of steps taken each day
```

```{r}
mean(totalSum) ## Mean of steps taken per day
median(totalSum) ## Median of steps taken per day
```

## What is the average daily activity pattern?

```{r}
mySortInd <- order(as.numeric(myProcData$interval), myProcData$interval)
mySortData <- myProcData[mySortInd,]

ii <- 2:length(mySortInd)
tempInterval <- mySortData$interval[1]
tempSum <- mySortData$steps[1]
cc <- 1
averageVector <- matrix()
intervalVector <- matrix()
j <- 1
for (i in ii)
{
  currentInterval <- mySortData$interval[i]
  if(currentInterval == tempInterval)
  {
    tempSum <- tempSum + mySortData$steps[i]
    cc <- cc + 1
  }
  else
  {
      averageVector[j] <- tempSum/cc ## Mean
      intervalVector[j] <- tempInterval
      j <- j + 1
      tempSum <- mySortData$steps[i]
      cc <- 1
      tempInterval <- mySortData$interval[i]
  }
}
averageVector[j] <- tempSum/cc ## Mean last 
intervalVector[j] <- tempInterval ## last Interval
```

```{r fig.width=7, fig.height=6}
plot(intervalVector,averageVector, type = "l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
intervalVector[averageVector == max(averageVector)]
```


## Imputing missing values
```{r}
sum(is.na(myData$steps) == "TRUE") ## Total number of missing values in the dataset

## Imputing by the the mean for that 5-minute interval

myNewData <- myData

dims <- dim(myNewData)

ii <- 1:dims[1]
for (i in ii)
{
  
  if (is.na(myData$steps[i]))
    {
      myNewData$steps[i] <- averageVector[intervalVector == myData$interval[i]]
    }
}

## Histogram of myNewData
ii <- 2:length(myNewData$steps)
tempDate <- myNewData$date[1]
tempSum <- myNewData$steps[1]
totalSum <- matrix()
j <- 1
for (i in ii)
{
  currentDate <- myNewData$date[i]
  if(currentDate == tempDate)
  {
    tempSum <- tempSum + myNewData$steps[i]
  }
  else
  {
      totalSum[j] <- tempSum
      j <- j + 1
      tempSum <- myNewData$steps[i]
      tempDate <- myNewData$date[i]
  }
}
totalSum[j] <- tempSum ## total sum of the last analyzed data
```

```{r fig.width=7, fig.height=6}
hist(totalSum) ## Histogram of the total number of steps taken each day
```

```{r}
## Slight changes in mean and median
mean(totalSum) ## Mean of steps taken per day
median(totalSum) ## Median of steps taken per day
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
myNewData$day <- as.POSIXlt(myNewData$date)$wday

WE <- myNewData$day != 6 & myNewData$day != 0

myNewData$WE <- WE

mySortInd <- order(as.numeric(myNewData$interval), myNewData$interval)
mySortData <- myNewData[mySortInd,]

ii <- 2:length(mySortInd)
tempInterval <-  mySortData$interval[1]
tempAve <- matrix(,1,2)
averageVector <- matrix(,1,2)
intervalVector <- matrix(,1,2)
j <- 1
for (i in ii)
{
  currentInterval <- mySortData$interval[i]
  if(currentInterval != tempInterval){
    test <- mySortData[mySortData$interval==tempInterval,]
    tempAve[1,1] <-sum(test$steps[test$WE=="TRUE"])/sum(test$WE=="TRUE")
    tempAve[1,2] <-sum(test$steps[test$WE=="FALSE"])/sum(test$WE=="FALSE")
    averageVector <- rbind(averageVector,tempAve)
    intervalVector[j] <- tempInterval
    j <- j+1
    tempInterval <- currentInterval
  }
}
intervalVector[j] <- tempInterval ## last Interval
test <- mySortData[mySortData$interval==tempInterval,]
tempAve[1,1] <-sum(test$steps[test$WE=="TRUE"])/sum(test$WE=="TRUE")
tempAve[1,2] <-sum(test$steps[test$WE=="FALSE"])/sum(test$WE=="FALSE")
averageVector <- rbind(averageVector,tempAve)
averageVector <- averageVector[2:(length(intervalVector)+1),]
```

```{r fig.width=7, fig.height=6}
plot(intervalVector,averageVector[,1], main = "Weekdays", type = "l")
```

```{r fig.width=7, fig.height=6}
plot(intervalVector,averageVector[,2], main = "Weekends", type = "l")
```
